---
title: "Intro datascience project"
output: html_document
date: "2023-10-10"
---

```{r}
# difficulty to find information over a significant period (example : survey/intentions data)
# strategy : information reliable to all the need of a family and factor related to population trend
# house/money/live_cost/   help/welfare?
# education/work
# sentiment : confiance/anxiety regardless to futur
# to hard : family policies, not really quantifiable(true??)
# other   : compare to other country.?????

# difficulty to find growth rate who are not just linear (otherwise it can easily follow the same trend without real correlation)
rm(list=ls())
data = read.csv("Data/data_final.csv",dec=",")
years = data[1]
fertility_rate = data[2]
```
difficulty to find information over a significant period (example : survey/intentions data)
strategy : information reliable to all the need of a family and factor related to population trend
house/money/live_cost/   help/welfare?
education/work
sentiment : confiance/anxiety regardless to futur
to hard : family policies, not really quantifiable(true??)
other   : compare to other country.?????

difficulty to find growth rate who are not just linear (otherwise it can easily follow the same trend without real correlation)
```{r}
plot(data[1:2])
```

```{r}
colnames(data)
```
```{r}
educations
```


```{r}
# Metric really significant because the dataset is small (32 rows)
to_drop = c()
data.cor = cor(data)[,2]
for (i in 1:length(data)){
  if(abs(data.cor[i])<0.15){
    to_drop = c(to_drop,i)
  }
}
colnames(data)[to_drop]

data2 = data[-c(to_drop)]
```
having basic educ do not matter "logic"
Basic education for 20+ : not important because pretty much every body have done it
Upper secondary education for fem20+/male30+ : same I suppose
Bachelor's or equivalent level for 15-19 : impossible/exceptional
Master's or equivalent level for 15-19 : impossible/exceptional
Doctoral or equivalent level for 15-19 year : impossible/exceptional

pop fem 25-39 not corr? => because fertility rate is already proportionate to pop size
risk of poverty of 0-17 & 30-49 not a prob? => more correlated to 18-29 which can be an age with more unstable situation (just after education)
shitty indicator : Consumer.confidence.indicator, Favourability.of.time.for.purchasing.durables

```{r}
educ.col = 47:129
colnames(data2[,educ.col])
data2.educ = data2[,educ.col]
```


```{r}
library(factoextra)
library(FactoMineR)
data2.educ.scale = scale(data2.educ)
res_pca = PCA(data2.educ.scale,graph = FALSE)
round(res_pca$eig,2)
```

the first component explain a lot 
we can try to keep as many as possible but to explain around 95% of the variance but it can be harder to the sharp decrease in variance of the components. (example : comp 4 explain only 3.8% of the total data variance)

```{r}
educations = c("Basic education", "3 Upper secondary education","4 Post-secondary non-tertiary education","5 Short-cycle tertiary education","6 Bachelor's or equivalent level","7 Master's or equivalent level","8 Doctoral or equivalent level")
```

```{r}
# options(ggrepel.max.overlaps = Inf)
# par( mar = c(4,10,4,1) )
# plot(res_pca,choix="var",select="cos2 0.92",axes=c(1,2),cex=0.7)+
#   scale_x_continuous(expand=c(0.5,0.5))
```

```{r}
n = 60
dim = 1

cos2.dim = res_pca$var$cos2[,dim]
var.dim <- names(sort(cos2.dim, decreasing = TRUE)[1:n])
coord.dim = res_pca$var$coord[var.dim,dim]

#####GRAPH######
best_var = order(cos2.dim,decreasing = TRUE)[1:n]
options(ggrepel.max.overlaps = Inf)
par( mar = c(4,10,4,1) )
plot(res_pca,choix="var",select=best_var,axes=c(1,2),cex=0.7)+
  scale_x_continuous(expand=c(0.5,0.5))

# dev.copy(png, "component1.png", width = 1200, height = 700, res = 200)
# dev.off()
```

```{r}
names(coord.dim[coord.dim>0])
```
positive dim1 :
Post-secondary non-tertiary education : male-female 25+
Bachelor's or equivalent level : male-female 20+
Master's or equivalent level : male-female 25+
Doctoral or equivalent level : male-female 30+

```{r}
names(coord.dim[coord.dim<0])
```

negative dim1 :
Basic education : 35+
Upper secondary education : male 15-19 | female 35-44 
Short-cycle tertiary education : fem 15+ | male 20+ (maybe related to short ending of education)
Bachelor's or equivalent level : male 20-24 (maybe military service)
Doctoral or equivalent level : male 25-29

Interpretation : 
High education vs short education + (military service for young men)

```{r}
thresh = 0.3
dim = 2

cos2.dim = res_pca$var$cos2[,dim]
var.dim <- names(sort(cos2.dim[cos2.dim>thresh], decreasing = TRUE))
coord.dim = res_pca$var$coord[var.dim,dim]

#####GRAPH######
best_var = which(cos2.dim>thresh)
options(ggrepel.max.overlaps = Inf)
par( mar = c(4,10,4,1) )
plot(res_pca,choix="var",select=best_var,axes=c(1,2),cex=0.7)+
  scale_x_continuous(expand=c(0.5,0.5))

dev.copy(png, "component2.png", width = 1200, height = 700, res = 200)
dev.off()
```
```{r}
coord.dim[coord.dim>0]
```
```{r}
coord.dim[coord.dim<0]
```

high dim 2: 
Basic educ : male 15-19
Upper secondary education : all 45+
Short-cycle tertiary education : female 40+ | male 45+
Master's or equivalent level : female 20-24
Doctoral or equivalent level : female 25-29 | male 30-34 


low dim 2:
Basic education : male 25-29 35+ | female 35+
Upper secondary education : male 25-29
Short-cycle tertiary education : all 15-19

Interpretation ~17% variation:
high education of youth (without big interruption) and short cycle education 40+ vs short cycle education of youth


```{r}
thresh = 0.2
dim = 3

cos2.dim = res_pca$var$cos2[,dim]
var.dim <- names(sort(cos2.dim[cos2.dim>thresh], decreasing = TRUE))
coord.dim = res_pca$var$coord[var.dim,dim]

#####GRAPH######
best_var = which(cos2.dim>thresh)
options(ggrepel.max.overlaps = Inf)
par( mar = c(4,10,4,1) )
plot(res_pca,choix="var",select=best_var,axes=c(1,3),cex=0.7)+
  scale_x_continuous(expand=c(0.5,0.5))

# dev.copy(png, "component3.png", width = 1200, height = 700, res = 200)
# dev.off()
```
```{r}
coord.dim[coord.dim>0]
```
```{r}
coord.dim[coord.dim<0]
```

High dim3 :
Basic education : female 20-24 | male 25-29

low dim 3:
Bachelor's or equivalent level : male 45-49

Interpretation (even if it explain only ~6 %) :
Basic educ 20-30 vs male bachelor 45+ (not really usefull.. and variable are not even well represented on this axis)

=> keep 2 component : 84% variation (not bad to compress 83 col)



```{r}
# New dataset with the 2 new component for all the education colones
data3 = data2[-educ.col]
data3["Educ.comp1"] = res_pca$ind$coord[,1]
data3["Educ.comp2"] = res_pca$ind$coord[,2]
data3.scale = scale(data3)
data3.df = as.data.frame(data3.scale)
# for (col in colnames(data3.df)){
#   data3.df[,col] = round(data3.df[,col],3)
# }
```

```{r}
# detach(data3.df)
# attach(data3.df)
```


```{r}
library(MASS)
full.model = lm(Total.fertility.rate ~ .-year, data = data3.df, na.action =  na.exclude)
summary(full.model)
#stepAIC(full.model, direction="both", trace=TRUE, k=2) # can't work because NA
```
Problem high dimension :
NA in the linear regression because there is to much correlation between var
Can be resolved by using PCA(dimensional reduction) but we don't want to loose all the data in 1 or 2 component, hard to interpret 


```{r}
library(glmnet)
```
```{r}
size = dim(data3.scale)
M = matrix(data=data3.scale, nrow = size[1], ncol=size[2])
colnames(M) = colnames(data3.scale)
```

```{r}
glmnet(x = M[,-c(1,2)], y = M[,2], alpha = 1, k=200)
```

```{r}
set.seed(27)

#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(x = M[,-c(1,2)], y = M[,2], alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda

#produce plot of test MSE by lambda value
plot(cv_model) 
```

```{r}
best_model <- glmnet(x = M[,-c(1,2)], y = M[,2], alpha = 1, lambda = 0.0001)
coef(best_model)
```

```{r}
lasso.coef = coef(best_model)
best_order = order(abs(lasso.coef), decreasing = TRUE)
row.names(lasso.coef)[best_order[1:length(lasso.coef[abs(lasso.coef)>0.1])]]
round(lasso.coef[best_order[1:length(lasso.coef[abs(lasso.coef)>0.1])]],2)
```
```{r}
"Pop.tot"                           
"Births.tot"                       
"Consumer.price.indice"             
"Pop.fem.15.19"                    
"Births.mother.20.24"               
"Births.mother.25.29"              
"Births.mother.0.19"                
"riskpoverty.rate.cashIncome.30.49"
"Death.crude.oo"                    
"Pop.mean"                         
"D1.Financial.situation.now"        
"Educ.comp2"                       
"B1.Own.economy.now..balance."      
"Deaths.infant.oo"                 
"Pop.fem.20.24"                     
"riskpoverty.rate.cashIncome.0.17" 
"riskpoverty.rate.cashIncome.18.29"


-0.58  0.38  0.24  0.23  0.18  0.17 -0.16  0.15 -0.15  0.14  0.14  0.13 -0.12 -0.12  0.12 -0.10  0.10
```

```{r}
plot(abs(lasso.coef))
```
```{r}
colnames(data)
```

# Graph with important feature


# futur step for the project

